---
title: "Billboard versus Grammy "
author: "Brittany"
date: "1/4/2020"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(timeDate)
library(ggplot2)
```


Read in grammy and billboard dataset - Grammys range from 1959-2019, Billboard range from August 1958 to June 2019. 
```{r}
grammy <- read_csv("~/Documents/nss_data_science/billboard-vs-grammy/data/grammy.csv")
billboard <- read_csv("~/Documents/nss_data_science/billboard-vs-grammy/data/billboard.csv")
```

Keep only Grammy General Fields (Big Four)
```{r}
grammy = filter(grammy, (category %in% c("Record Of The Year", "Album Of The Year","Song Of The Year","Best New Artist")))  %>%
  rename(Performer = 'name', Song = 'awardFor') %>%
  select(-c('X5','X6'))
```

Remove columns from Billboard
```{r}
billboard <- billboard %>% 
  select(-c('SongID','Peak Position', 'Weeks on Chart', 'X10', 'X11','X12','X13','X14','X15','X16','X17','X18'))
```

Create "position difference" column with Billboard position changes
```{r}
billboard$'Week Position Difference' <- billboard$'Week Position' - billboard$'Previous Week Position'
```

To fix NA values in the merged dataset, edit the Performer name to match in merge
```{r}
billboard <- billboard %>%
  mutate(Performer = str_replace(Performer, "Macklemore & Ryan Lewis Featuring Eric Nally, Melle Mel, Kool Moe Dee & Grandmaster Caz","Macklemore & Ryan Lewis")) %>% 
  mutate(Performer = str_replace(Performer, "Macklemore & Ryan Lewis Featuring Mary Lambert","Macklemore & Ryan Lewis")) %>%
  mutate(Performer = str_replace(Performer, "Macklemore & Ryan Lewis Featuring Ray Dalton", "Macklemore & Ryan Lewis")) %>%
  mutate(Performer = str_replace(Performer, "Macklemore & Ryan Lewis Featuring ScHoolboy Q & Hollis", "Macklemore & Ryan Lewis")) %>%
  mutate(Performer = str_replace(Performer, "Macklemore & Ryan Lewis Featuring Wanz", "Macklemore & Ryan Lewis"))
```

Create "year" column in Grammy dataset
```{r}
grammy$year <- 1958 + grammy$annualGrammy
```

Convert Billboard "WeekID" column to date
```{r}
billboard$WeekID <- as.Date(billboard$WeekID , "%m/%d/%Y")
```

Create Grammy_artist dataset for Best New Artist category
```{r}
grammy_artist <- filter(grammy, (category %in% ("Best New Artist")))
```

Create Grammy_music dataset for Record of the Year and Song of the Year categories
```{r}
grammy_song <- filter(grammy, (category %in% c("Record Of The Year","Song Of The Year")))
```

Create Grammy_album dataset for Album of the Year categories
```{r}
grammy_album <- filter(grammy, (category %in% c("Album Of The Year")))
```

Merge Billboard with grammy_music
```{r}
song_merge <- merge(x=grammy_song,y=billboard,by=c("Performer","Song"),all.x=TRUE)
```

Merge Billboard with grammy_artist
```{r}
artist_merge <- merge(x=grammy_artist,y=billboard,by="Performer",all.x=TRUE)
```

Merge Billboard with grammy_album
```{r}
album_merge <- merge(x=grammy_album,y=billboard,by="Performer",all.x=TRUE)
```

Create billboard_album with billboard hits and album name
```{r}
billboard_album <- billboard
```
Add empty column to billboard_album to add album names
```{r}
billboard_album$album <- NA
```

Fill album rows
```{r}
fill_album <- function(df, song, album_name){
  df %>% mutate(album = if_else(Song == song, album_name, album))
}
```  

```{r}
billboard_album %>% fill_album('By The Time I Get To Phoenix', 'By The Time I Get To Phoenix') %>% View()

```

```{r}
billboard_album$album <- ifelse(billboard_album$Song == 'By The Time I Get To Phoenix', 'By The Time I Get To Phoenix', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Hey Little One', 'By The Time I Get To Phoenix', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'So Far Away/Smackwater Jack', 'Tapestry', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Bangla-Desh/Deep Blue', 'The Concert For Bangla Desh', NA)
billboard_album$album <- ifelse(billboard_album$Song == "My Sweet Lord/Isn't It A Pity", 'The Concert For Bangla Desh', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Higher Ground', 'Innervisions', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Living For The City', 'Innervisions', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Boogie On Reggae Woman', "Fulfillingness' First Finale", NA)	
billboard_album$album <- ifelse(billboard_album$Song == "You Haven't Done Nothin", "Fulfillingness' First Finale", NA)
billboard_album$album <- ifelse(billboard_album$Song == '50 Ways To Leave Your Lover', 'Still Crazy After All These Years', NA)
billboard_album$album <- ifelse(billboard_album$Song == "Slip Slidin' Away", 'Still Crazy After All These Years', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'I Wish', 'Songs In The Key Of Life', NA)
billboard_album$album <- ifelse(billboard_album$Song == "Sir Duke", 'Songs In The Key Of Life', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'It Was A Very Good Year', 'Innervisions', NA)
billboard_album$album <- ifelse(billboard_album$Song == "Sgt. Pepper's Lonely Hearts Club Band/With A Little Help From My Friends", "Sgt. Pepper's Lonely Hearts Club Band", NA)
billboard_album$album <- ifelse(billboard_album$Song == 'My Way', "A Man And His Music", NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Still Crazy After All These Years', 'Still Crazy After All These Years', NA)
billboard_album$album <- ifelse(billboard_album$Song == "Say You'll Be Mine", 'Christopher Cross', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Never Be The Same', 'Christopher Cross', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Sailing', 'Christopher Cross', NA)
billboard_album$album <- ifelse(billboard_album$Song == "Ride Like The Wind", 'Christopher Cross', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Watching The Wheels', 'Double Fantasy', NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Woman', 'Double Fantasy', NA)
billboard_album$album <- ifelse(billboard_album$Song == '(Just Like) Starting Over', "Double Fantasy", NA)	
billboard_album$album <- ifelse(billboard_album$Song == "Thriller", "Thriller", NA)
billboard_album$album <- ifelse(billboard_album$Song == 'Billie Jean', 'Thriller', NA)
billboard_album$album <- ifelse(billboard_album$Song == "P.Y.T. (Pretty Young Thing)", 'Thriller', NA)
```
```{r}
billboard_album$album <- billboard_album %>%
  mutateifelse(billboard_album$Song == 'Billie Jean', 'Thriller', NA)
```

Remove annualGrammy from grammy dataset
```{r}
grammy <- grammy %>% select(-c('annualGrammy'))
```

Test filter
```{r}
grammy_filter <- grammy %>%
            filter(year == 2019)
```

Remove duplicate Song of the Year row
```{r}
grammy <- grammy[-c(39,40,41),]
```

Time series example
```{r}
filteredSong <- song_merge %>%
  filter(year == 2019) %>%
  filter(category == "Song Of The Year")
```

```{r}
qplot(x=filteredSong$WeekID, y=filteredSong$`Week Position`,
    main="Billboard Chart Movement",
    xlab="Week", ylab="Week Position")
```